---
title: "final project on exercise"
output: html_document
---
BACKGROUND AND INSTRUCTIONS

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


DATA SOURCES 

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

DATA Analysis 

Load the data first 

```{r setup, include=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(randomForest)
library(party)
```

set seed to reproduce
```{r cars}
set.seed(12345)
```

Then let's get the data 

```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl),na.strings = c("NA",""))
testing <- read.csv(url(testUrl),na.strings = c("NA",""))
```

Then briefly look at the number of observations in each dataset 

```{r}
nrow(training)
ncol(training)
nrow(testing)
ncol(testing)
```

Data partitioning where 60/40 is for training/testing 
```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]
myTesting <- training[-inTrain, ]
```

Let's try both decision trees and random forest testing. Because it takes a long time to run, let's take out near zero variance variables. I also removed columns with NA values (You can choose to keep them but my laptop is really slow so I have to this choice)

```{r}
NZVdata <- nearZeroVar(myTraining, saveMetrics=TRUE)
str(NZVdata, vec.len=2)
NZVdata[NZVdata[,"zeroVar"] + NZVdata[,"nzv"] > 0, ] 
```

```{r}
nzv_cols <- nearZeroVar(myTraining)
if(length(nzv_cols) > 0) myTraining <- myTraining[, -nzv_cols]
```

Remove NA values so the final columns are 59 instead of the original 160. 
```{r}
train_clean<-myTraining[, colSums(is.na(myTraining)) == 0]
test_clean<-myTesting[, colSums(is.na(myTesting)) == 0]
dim (train_clean)
```


Then running tree graph with 5-fold validation 

```{r}
control <- trainControl(method = "cv", number = 5)
model_ctree <- train(classe ~ ., data = train_clean, method = "ctree", trControl = control)
print(model_ctree, digits = 5)
```

Then we make predictions on the test set
```{r}
predict_ctree <- predict(model_ctree, test_clean)
confusionMatrix(test_clean$classe, predict_ctree)
```

The accuracy of 0.9997 means out of sample error is 0.0003. 

Next similarily, we try random forest 

```{r}
model_rf <- train(classe ~ ., data = train_clean, method = "rf", trControl = control)
print(model_rf, digits = 5)
predict_rf <- predict(model_rf, test_clean)
confusionMatrix(test_clean$classe, predict_rf)
```

The accuracy is 0.9999, with error of 0.0001 so it beats decision trees. So we will use this method as our final model. 

```

